### 2017-07-14 Plos-one-manuscript/latex/manuscript-old.tex
ready to submit

One example of an everyday scientific tool that provides opportunity for an open design solution is the micropipette.
An open design micropipette that can be made cheaply affords more options for labs and educational settings.
Micropipettes are an indispensable tool used routinely in lab tasks and can easily cost \$1000 USD for a set.
Often a lab will require several sets each for a dedicated task.
Some pipettes may even be re-calibrated for use with liquids of different properties.

Air displacement pipettes use a piston operating principle to draw liquid into the pipette\cite{ISO8655}.
In a typical commercial pipette the piston is made to be gas-tight with a gasketed plunger inside a smooth barrel.
Consumer grade fused deposition modeling (FDM) printers are unable to build a smooth surface due to the formation of ridges that occur as each layer is deposited\cite{Takagishi2017}.
The ridges formed by FDM make it impractical to form a gas-tight seal between moving parts, even with a gasket.
Existing printable open-design micropipettes get around this limitation by stretching a membrane over one end of a printed tube, which when pressed causes the displacement.
The displacement membrane can be made from any elastic material such as a latex glove.
A few open design micropipettes exist including a popular one which, in addition to the printed parts, uses parts scavenged from a retractable pen\cite{Baden2014}.
Because there is no built in feature such as a readout for the user to set the displacement to a desired volume, this design requires the user to validate the volumes dispensed with a high precision scale.
Without verification with a scale the volumes dispensed can only be estimated based on calculations of the deflection of the membrane which is not a practical protocol.

@@

\section*{Methods}
Our printed pipette is designed to actuate a 1 mL or 3 mL syringe to a user set displacement.
The core of the design is two printed parts, the body and the plunger which are able to be printed on a consumer grade FDM printer (fig.~\ref{fig1}), in this case a Makerbot Replicator.
A 1 mL or 3 mL syringe twists to lock in the body part and is held into place by the syringe flanges.
The 30-300 $\mu$L configuration uses a 1 mL pipette and the 100-1000 $\mu$L configuration uses a 3 mL pipette.
The plunger part slides freely in the body part and actuates the syringe by pushing the thumb button (fig.~\ref{fig2}). 
The pipette is spring loaded towards a set point which is adjustable by a set-screw.
When the thumb button is pressed the system locks when it reaches the latched position, where it is ready to draw in fluid.
The plunger is held in place with a latching button design which is released with the unlatching button drawing in fluid.
The displacement is equal to the distance between the set position and the latched position.
The pipette can also be pressed past the latched position to `blow-out' the transferred fluid completely from the pipette tip.   
Additional materials required for assembly include two springs, a nut, and a bolt.
Attempts to make a printable luer lock adapter for tips was abandoned as the surface of printed parts is too rough to make an air tight seal with the luer or pipette tip.
Instead, a combination of a barbed luer adapter and elastic tubing is used to attach the pipette tips.
Our printed pipette mimics commercial pipettes design, function, and user operation, making it intuitive to use.

@@

\begin{figure}
	\caption{
		{\bf CAD renderings of assembled pipette and function.} The pipette actuates the syringe to three positions. {\bf (A) The latched position.} When the plunger is pressed the pipette locks at this position. The tip is then placed in a liquid and the unlatching button is pressed to release the pipette back to the set position (B), drawing in liquid.  {\bf(B) The set position.} The position of the screw determines the total displacement that the plunger moves. The pipette is spring loaded to return to this position. {\bf (C) The blow-out position.} The fluid is transferred by pressing the plunger past the latched position to blow-out all the liquid. {\bf (D) Return to the latched position.} The pipette returns to the latched position ready to perform another transfer.
	}
	\label{fig2}
\end{figure}

@@

\subsection*{Fabrication \& Assembly}
Two parts are printed (body.stl and plunger.stl) at normal resolution with rafts on a Makerbot Replicator.
A small amount of paraffin wax is applied to the screw to prevent slop from causing the set point to drift after each actuation.
The nut is sunk into the hex inset in the printed body part.
The bolt is threaded in from the top of the body into the nut.
Two springs are threaded onto the plunger of a 1 mL syringe for the 30-300 $\mu$L configuration.
Springs are placed inside a 3 mL syringe for the 100-1000 $\mu$L configuration.
The plunger part is inserted in the body and the syringe assembly is pushed in and locked from the syringe flanges to the body part to complete assembly(fig.~\ref{fig3}).
Parts and cost are listed in Table~\ref{table1} and \ref{table2}.

@@

\begin{table}[!ht]
\centering
\caption{Parts and cost for the 30-300 $\mu$L pipette.}
\label{table1}
	\begin{tabular}{|l|l|r|l|}
		\hline
	30-300 mL Configuration:              & Source            & Unit Price & Part Number \\ \hline
	Filament                              & Makerbot          & \$1.63     & NA          \\ \hline
	1 mL Syringe                          & BD Biosciences    & \$0.15     & 309628      \\ \hline
	M3 Bolt, 35 mm                        & McMaster-Carr     & \$0.12     & 91287A026   \\ \hline
	M3 Nut                                & McMaster-Carr     & \$0.01     & 90591A121   \\ \hline
	Music Wire Compression Springs* (2)   & Jones Spring Co   & \$1.23     & C10-022-048 \\ \hline
	Female Luer to 1/8" Hose Barb Adapter & Cole-Parmer       & \$0.40     & EW-30800-08 \\ \hline
	Tygon Tubing, 1/16"ID x 3/16"OD       & Cole-Parmer       & \$0.03     & EW-06407-72 \\ \hline
	300 uL Tips (10)                      & Fisher Scientific & \$0.34     & 02-707-447  \\ \hline
	& Total             & \$3.91     &     \\ \hline        
	\end{tabular}
\end{table}

\begin{table}[!ht]
\centering
\caption{Parts and cost for the 100-1000 $\mu$L pipette.}
\label{table2}
\begin{tabular}{|l|l|r|l|}
	\hline
	100-1000 mL configuration:             & Source            & Unit Price & Part Number \\ \hline
	Filament                               & Makerbot          & \$1.63     & NA          \\ \hline
	3 mL Syringe                           & BD Biosciences    & \$0.73     & 309657      \\ \hline
	M3 Bolt, 35 mm                         & McMaster-Carr     & \$0.12     & 91287A026   \\ \hline
	M3 Nut                                 & McMaster-Carr     & \$0.01     & 90591A121   \\ \hline
	Music Wire Compression Springs* (2)    & Jones Spring Co   & \$1.23     & C10-022-048 \\ \hline
	Female luer to 5/32" Hose Barb Adapter & Cole-Parmer       & \$0.66     & EW-45508-06 \\ \hline
	Tygon Tubing, 1/8"ID x 1/4"OD          & Cole-Parmer       & \$0.05     & EW-06407-76 \\ \hline
	1000 uL Tips (10)                      & Fisher Scientific & \$0.41     & 02-707-400  \\ \hline
	& Total             & \$4.84     &            \\ \hline
\end{tabular}
\begin{flushleft} *Spring Specifications: Overall length 38 mm, OD 7.62 mm, wire diameter 0.56 mm, Max load 0.86 kg, with closed and flat ends. Also available from McMaster-Carr P/N: 9657K347.
\end{flushleft}
\end{table}

@@

\section*{Results and Discussion}
Validation testing performed with the existing graduations printed on the syringes did not meet the ISO standards.
According to ISO 8655 for a pipette with the maximum nominal volume of 1000 $\mu$L the systematic error cannot exceed 8 $\mu$L and the random error cannot exceed 3 $\mu$L.
For the maximum nominal volume of 300 $\mu$L the systemic error cannot exceed 4 $\mu$L and the random error cannot exceed 1.5 $\mu$L.
The commercial pipettes met these standards handily, but in our initial tests with our printed pipette we noticed large negative systematic error suggesting that we were missing a biasing factor. 
This error was exaggerated while transferring larger volumes.
After further investigation we realized that the graduations are intended to measure incompressible fluids within the syringe while we were using them to measure air while under vacuum.
Because the syringe graduations were used to measure the air under vacuum it could not be expected to serve as a 1:1 volume ratio of water that was pulled into the tip because under vacuum the air had expanded.
This explains the negative systematic error we were experiencing.
The water in the tip pulls on the air due to gravity which causes it to expand slightly but enough to make measuring with the graduation in this way inaccurate.

Our solution was to replace the built-in scale with a scale that accounted for the expansion.
We compared expected volumes with actual measurements and noticed a linear relationship.
We scaled up the graduations by a factor of 1.027x for both the 100-1000 $\mu$L, and 30-300 $\mu$L configurations.
These scales were exported to portable document format (PDF) so that they are most easily printed to scale.
These adjusted scales are printed on paper or a transparency sheet and taped onto the syringe with the zero at the plunger in the latched position (\nameref{S1_Fig} and \nameref{S1_Fig}).

With the new scale taped on the syringe our validation testing met the ISO 8655 standard.
Replacing the scale with our scale that accounts for the expansion corrected this effect. 
Our printed pipette with the adjusted scale meets ISO standards for accuracy and precision. (Table \ref{table3} and \ref{table4}).

@@

Our printed pipette improves on existing open design pipettes in several ways. 
Most significantly the user is able to adjust the syringe accurately without verifying the volume with a scale.
The pipette can be adjusted to any discrete volume within range.
In addition, the assembly requires no permanent connections using tape or glue which allows for re-configuration and easy replacement of parts.
Conveniently, our design can also reach to the bottom of a 15 mL conical tube allowing tasks such as aspirating supernatant fluid from a cell pellet.
Unlike existing designs our design does not use a membrane that may wear out and require replacement.
The only major limitation of this design compared to the biropipette is the option for a pipette tip ejector.
The biropipette also uses arguably easier to source parts (a pen VS a syringe) relative to what is required from our design.
The biropipette also was developed in OpenSCAD which is free and open source CAD software where our design was created in Solidworks a proprietary software.
This creates a limitation for users who may not have access to the proprietary software used to make the design.
Fortunately the produced STL files are universal and can be imported an modified by any CAD software.

The luer to pipette tip connection is a good candidate for a 3D printed solution.
Unfortunately the current state of FDM can not make this adapter with high enough resolution or critically, surface smoothness.
Other printing techniques, such as stereolithography (SLA) and MultiJet Printing (MJP), provide higher resolution and smoother finishing, yet FDM is the the most widely available and affordable of printing technologies.
SLA printers are coming down in price but are still significantly more expensive especially considering the higher cost of the consumable resin vs. FDM filament.
Ultimately the goal of this project is to make this pipette cheap and easy to make.

@@

The open design nature of this project encourages anyone to use and submit changes to improve and add features to this design.
All of the working files and documentation are kept in a public repository~\cite{BMLmicropipette}.
Future directions for this project include developing a tip ejection system and additional configurations for transferring volumes below 30 uL as well as increasing over all user friendliness, such as making the adjusting bolt easier to grip and improving ergonomics.
It's our hope that this project is under continuous development and compatibility with different syringes and printers can be crowd-sourced from community contributors.

@@

\subsection*{S1 Video}
\label{S1_Video}
{\bf Video showing assembly and demonstration of operation.} Assembly of the pipette in the 100-1000 $\mu$L and 30-300 $\mu$L configuration is performed. https://www.youtube.com/watch?v=D8y5WUYpcW0

>>>>

One example of a everyday scientific tool that provides opportunity for an open design solution is the micropipette.
An open design micropipette that can be made cheaply affords more options for labs and educational settings.
Micropipettes are an indispensable tool used routinely in lab tasks and can easily cost \$1000 USD for a set.
Often a lab will require several sets each for a dedicated task.
Some pipettes may even re-calibrated for use with liquids of different properties.

Air displacement pipettes use a piston operating principal to draw liquid into the pipette\cite{ISO8655}.
In a typical commercial pipette the piston is made to be gas-tight with a gasketed plunger inside a smooth barrel.
Consumer grade fused deposition modeling (FDM) printers are unable to build a smooth surface do to the formation of ridges that occur as each layer is deposited\cite{Takagishi2017}.
The ridges formed by FDM make it impractical to form a gas-tight seal between moving parts, even with a gasket.
Existing printable open-design micropipettes get around this limitation by stretching a membrane over one end of a printed tube, which when pressed causes the displacement.
The displacement membrane can be made from any elastic material such as a latex glove.
A few open design micropipettes exist including a popular one which, in addition to the printed parts, uses parts scavenged from a retractable pen\cite{Baden2014}.
Because there is no built in feature such as a readout for the user to set the displacement to a desired volume, this design requires the user to validate the volumes dispensed with a high precision scale.
Without verification with a scale the volumes dispensed can only be estimated based on calculations of the deflection of the membrane which is not a practical protocol.

@@

\section*{Methods}
Our printed pipette is designed to actuate a 1 mL or 3 mL syringe to a user set displacement.
The core of the design is two printed parts, the body and the plunger which are able to be printed on a consumer grade FDM printer (fig.~\ref{fig1}), in this case a Makerbot Replicator.
A 1 mL or 3 mL syringe twists to lock in the body part and is held into place by the syringe flanges.
The 30-300 $\mu$L configuration uses a 1 mL pipette and the 100-1000 $\mu$L configuration uses a 3 mL pipette.
The plunger part slides freely in the body part and actuates the syringe by pushing the thumb button (fig.~\ref{fig2}). 
The pipette is spring loaded towards a set point which is adjustable by a set-screw.
When the thumb button is pressed the system locks when it reaches the latched position, where it is ready to draw in fluid.
The plunger is held in place with a latching button design which is released with the unlatching button drawing in fluid.
The displacement is equal to the distance between the set position and the latched position.
The pipette can also be pressed past the latched position to `blow-out' the transferred fluid completely from the pipette tip.   
Additional materials required for assembly include two springs, a nut, a bolt and two washers.
Attempts to make a printable luer lock adapter for tips was abandoned as the surface of printed parts is too rough to make an air tight seal with the luer or pipette tip.
Instead, a combination of a barbed luer adapter and elastic tubing is used to attach the pipette tips.
Our printed pipette mimics commercial pipettes design, function, and user operation, making it intuitive to use.

@@

\begin{figure}
	\caption{
		{\bf CAD renderings of assembled pipette and function.} The pipette actuates the syringe to three positions. {\bf (A) The latched position.} When the plunger is pressed the pipette locks at this position. The tip is then placed in a liquid and the unlatching button is pressed to release the pipette back to the set position (B), drawing in liquid.  {\bf(B) The set position.} The position of the screw determines the total displacement that the plunger moves. The pipette is spring loaded to return to this position. {\bf (C) The blow-out position.} The fluid is transferred by pressing the plunger past the latched position to blow-out all the liquid. {\bf (D) Return to the latched position.} The pipette returns to the latched position ready to preform another transfer.
	}
	\label{fig2}
\end{figure}

@@

\subsection*{Fabrication \& Assembly}
Two parts are printed (body.stl and plunger.stl) at normal resolution with rafts on a Makerbot Replicator.
A small amount of paraffin wax is applied to the screw to prevent slop from causing the set point to drift after each actuation.
The nut is sunk into the hex inset in the printed body part.
The bolt is threaded in from the top of the body into the nut.
Springs and washers are threaded onto the plunger of a 1 mL syringe for the 30-300 $\mu$L configuration.
Springs are placed inside a 3 mL syringe for the 100-1000 $\mu$L configuration.
The plunger part is inserted in the body and the syringe assembly is pushed in and locked from the syringe flanges to the body part to complete assembly(fig.~\ref{fig3}).
Parts and cost are listed in Table~\ref{table1} and \ref{table2}.

@@

\begin{table}[!ht]
\centering
\caption{Parts and cost for the 30-300 $\mu$L pipette.}
\label{table1}
\begin{tabular}{|l|l|l|l|}
	\hline
	Part         & Unit Price & Source         & Part number \\ \hline
	Filament     & \$1.63     & Makerbot       & NA          \\ \hline
	1 mL Syringe & \$0.15     & BD Biosciences & 309628      \\ \hline
	Bolt         & \$0.12     & McMaster-Carr  & 91287A026   \\ \hline
	Nut          & \$0.01     & McMaster-Carr  & 90591A121   \\ \hline
	Spring (2)   & \$4.14     & McMaster-Carr  & 94125K542   \\ \hline
	Washers (2)  & \$0.16     & McMaster-Carr  & 90107A012   \\ \hline
	total        & \$6.21     &                &             \\ \hline
\end{tabular}
\end{table}

\begin{table}[!ht]
\centering
\caption{Parts and cost for the 100-1000 $\mu$L pipette.}
\label{table2}
\begin{tabular}{|l|l|l|l|}
	\hline
	Part         & Unit Price & Source         & Part number \\ \hline
	Filament     & \$1.63     & Makerbot       & NA          \\ \hline
	3 mL Syringe & \$0.73     & BD Biosciences & 309657      \\ \hline
	Bolt         & \$0.12     & McMaster-Carr  & 91287A026   \\ \hline
	Nut          & \$0.01     & McMaster-Carr  & 90591A121   \\ \hline
	Spring (2)   & \$4.14     & McMaster-Carr  & 94125K542   \\ \hline
	total        & \$6.63     &                &             \\ \hline
\end{tabular}
\end{table}

@@

\section*{Results and Discussion}
Validation testing preformed with the existing graduations printed on the syringes did not meet the ISO standards.
According to ISO 8655 for a pipette with the maximum nominal volume of 1000 $\mu$L the systematic error cannot exceed 8 $\mu$L and the random error cannot exceed 3 $\mu$L.
For the maximum nominal volume of 300 $\mu$L the systemic error cannot exceed 4 $\mu$L and the random error cannot exceed 1.5 $\mu$L.
The commercial pipettes met these standards handily, but in our initial tests with our printed pipette we noticed large negative systematic error suggesting that we were missing a biasing factor. 
This error was exaggerated while transferring larger volumes.
After further investigation we realized that the graduations are intended to measure incompressible fluids within the syringe while we were using them to measure air while under vacuum.
Because the syringe graduations were used to measure the air under vacuum it could not be expected to serve as a 1:1 volume ratio of water that was pulled into the tip because under vacuum the air had expanded.
This explains the negative systematic error we were experiencing.
The water in the tip pulls on the air due to gravity which causes it to expand slightly but enough to make measuring with the graduation in this way inaccurate.
Our solution was to replace the built-in scale with a scale that accounted for the expansion.
Using the data from our initial testing we calculated a factor for the expansion and used that to make the new scale (\nameref{S1_Fig} and \nameref{S1_Fig}.)
With the new scale taped on the syringe our validation testing met the ISO 8655 standard.
Replacing the scale with our scale that accounts for the expansion corrected this effect. 
Our printed pipette with the adjusted scale meets ISO standards for accuracy and precision. (Table \ref{table3} and \ref{table4}).

@@

Our printed pipette improves on existing open design pipettes in several ways. 
Most significantly the user is able to adjust the syringe accurately without verifying the volume with a scale.
The pipette can be adjusted to any discrete volume within range.
In addition, the assembly requires no permanent connections using tape or glue which allows for re-configuration and easy replacement of parts.
Conveniently, our design can also reach to the bottom of a 15 mL conical tube allowing tasks such as aspirating supernatant fluid from a cell pellet.
Unlike existing designs our design does not use a membrane that may wear out and require replacement.
The only major limitation of this design compared to the biropipette is the option for a pipette tip ejector.
The biropipette also uses arguably easier to source parts (a pen VS a syringe) relative to what is required from our design.
The biropipette also was developed in OpenSCAD which is free and open source CAD software where our design was created in Solidworks a proprietary software.

@@

The open design nature of this project encourages anyone to use and submit changes to improve and add features to this design.
All of the working files and documentation are kept in a public repository~\cite{BMLmicropipette}.
Future directions for this project include developing a tip ejection system and additional configurations for transferring volumes below 30 uL as well as increasing over all user friendliness, such as making the adjusting bolt easier to grip and improving ergonomics.

@@

\subsection*{S1 Video}
\label{S1_Video}
{\bf Video showing assembly and demonstration of operation.} Assembly of the pipette in the 100-1000 $\mu$L and 30-300 $\mu$L configuration is preformed. https://www.youtube.com/watch?v=D8y5WUYpcW0
